---
title: "Regression using Categorical Variable"
author: "Ashutosh Kumar"
date: "2/22/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Some examples of using categorical variables.
```{r echo=TRUE}
library(ISLR)
data(Credit)
?Credit
```
## Response variable of interest is 'Balance'
## Predictor: 'student': Yes or No
```{r echo=TRUE}
Credit$Student

Credit$std_dummy = ifelse(Credit$Student == "Yes", 1,0)
mod1 = lm(Balance~std_dummy, data = Credit)
summary(mod1)

Credit$std_dummy = ifelse(Credit$Student == "Yes", 0,1)
mod2 = lm(Balance~std_dummy, data = Credit)
summary(mod2)

```
Without creating specific dummies, we can still do it.

```{r echo=TRUE}
mod2 = lm(Balance~ Student, data = Credit)
summary(mod2)
#model.matrix(mod2) this puts all data in matrix form
```
## Categorical variable with more than two levels
## The variable 'Ethnicity' has three levels
```{r echo=TRUE}
mod3 = lm(Balance~ relevel(Ethnicity, ref = "Asian"), data = Credit)
summary(mod3)
#model.matrix(mod3) this puts all data in matrix form
str(Credit)
```

Above we put Asian as the reference, instead of default given by R as 'African American'.

```{r echo=TRUE}
mod4 = lm(Balance~.-ID-std_dummy, data = Credit)
summary(mod4)
model.matrix(mod4)
```

# Example 2:

```{r echo=TRUE}
n=100
y=rnorm(n)
x = rep(c(1,2,3,4,5),n/5) #Assume that 1,2,3,4,5 represent qualitative levels
length(x)

mod5 = lm(y~as.factor(x))

mod.reduced = lm(y~1)
anova(mod.reduced, mod5) # Here we compared two models. full model and reduced model with null saying reduced is better.
summary(mod5)
```

# Example 3:

```{r echo=TRUE}
data("Carseats")
?Carseats

mod6 = lm(Sales~., data = Carseats)
summary(mod6)
```

To determine whether the categorical variable 'Shelveloc' is significantly associated with 'Sales'

```{r echo=TRUE}
mod.red = lm(Sales~.-ShelveLoc, data = Carseats)
av = anova(mod.red, mod6)
av$`Pr(>F)`
```
##Simulate Data
```{r echo=TRUE}
n = 100
x = rnorm(n)
y = exp(x)+rnorm(n)
plot(x,y)

mod7 = lm(y~x+I(x^2)+I(x^4)+I(x^5)+I(x^6))
summary (mod7)

mod8 = lm(y~x+I(x^2)+I(x^4)+I(x^5))
summary (mod8)

```